The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-07 manuscript/manuscript.Rnw
discuss horse placenta transcriptome

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \mouse are shown in
\cref{tab:top50_mouse_placenta} with a list of all genes with
$\mathrm{FPKM} \ge 50$ in Supplemental
\cref{tab:all_mus_musculus_placenta}. The top expressed genes include
\Sexpr{gene.expression("mus_musculus","Mir6236",complete=TRUE)}, whose
target \textit{in silico} is the orphan G-coupled receptor
\textit{Mrgprg},
\Sexpr{gene.expression("mus_musculus","H19",complete=TRUE)} and
\Sexpr{gene.expression("mus_musculus","Gm27786",complete=TRUE)}, both
of which are maternally imprinted (the later of which is a homolog of
human \textit{H19\_2}),
\Sexpr{gene.expression("mus_musculus","IGF2",complete=TRUE)}, numerous
cathepsins and a rodentia specific expansion of prolactin related
genes. There are also two miRNA which overlap with the of the 7S
signal recognition particle
(\Sexpr{gene.expression("mus_musculus","AC157822.1",complete=TRUE)}
and
\Sexpr{gene.expression("mus_musculus","AC099934.1",complete=TRUE)})
and numerous pseudogenes
(\Sexpr{gene.expression("mus_musculus","Gm11966",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10736",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10180",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm29216",complete=TRUE)}) which
appear to be highly expressed, however, the short read length (34--44)
in many samples is likely preventing unambiguous assignment of all
reads to the more lengthy complete transcript, so these shorter
transcripts should be treated with suspicion, and we will ignore them
in further discussion. The prolactin-related genes
(\Sexpr{gene.expression("mus_musculus","Prl3b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl2b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl8a9",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Prl8a6",complete=TRUE)}) are
very highly expressed in \mouse in, with %
\Sexpr{combined.fpkm[species=="mus musculus" &
  grepl("Prl(\\d.\\d)?",gene_short_name), sum(mean_fpkm >= 1)]}
prolactin related genes having $\mathrm{FPKM} \ge 1$
(\cref{fig:prolactin_evolution_tree}). The cathespins have also
expanded in rodents, with six showing high expression
(\Sexpr{gene.expression("mus_musculus","Ctsj",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts6",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsr",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsq",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts3",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Ctsm",complete=TRUE)},
\cref{tab:top50_mouse_placenta,fig:cathespin_evolution_tree}).
Finally, \textit{Peg10}, an imprinted gene required for proper
formation of the placenta in mice is also highly expressed
(\Sexpr{gene.expression("mus_musculus","Peg10")}), as previously
reported \cite{Ono.ea2006:DeletionofPeg10imprinted}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Equus Caballus} are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}.

>>>>

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \mouse are shown in
\cref{tab:top50_mouse_placenta} with a list of all genes with
$\mathrm{FPKM} \ge 50$ in Supplemental
\cref{tab:all_mus_musculus_placenta}. The top expressed genes include
\Sexpr{gene.expression("mus_musculus","Mir6236",complete=TRUE)}, whose
target \textit{in silico} is the orphan G-coupled receptor
\textit{Mrgprg}
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget},
\Sexpr{gene.expression("mus_musculus","H19",complete=TRUE)} and
\Sexpr{gene.expression("mus_musculus","Gm27786",complete=TRUE)}, both
of which are maternally imprinted (the later of which is a homolog of
human \textit{H19\_2}),
\Sexpr{gene.expression("mus_musculus","IGF2",complete=TRUE)}, numerous
cathepsins and a rodentia specific expansion of prolactin related
genes. There are also two miRNA which overlap with the of the 7S
signal recognition particle
(\Sexpr{gene.expression("mus_musculus","AC157822.1",complete=TRUE)}
and
\Sexpr{gene.expression("mus_musculus","AC099934.1",complete=TRUE)})
and numerous pseudogenes
(\Sexpr{gene.expression("mus_musculus","Gm11966",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10736",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10180",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm29216",complete=TRUE)}) which
appear to be highly expressed, however, the short read length (34--44)
in many samples is likely preventing unambiguous assignment of all
reads to the more lengthy complete transcript, so these shorter
transcripts should be treated with suspicion, and we will ignore them
in further discussion. The prolactin-related genes
(\Sexpr{gene.expression("mus_musculus","Prl3b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl2b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl8a9",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Prl8a6",complete=TRUE)}) are
very highly expressed in \mouse in, with %
\Sexpr{combined.fpkm[species=="mus musculus" &
  grepl("Prl(\\d.\\d)?",gene_short_name), sum(mean_fpkm >= 1)]}
prolactin related genes having $\mathrm{FPKM} \ge 1$
(\cref{fig:prolactin_evolution_tree}). The cathespins have also
expanded in rodents, with six showing high expression
(\Sexpr{gene.expression("mus_musculus","Ctsj",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts6",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsr",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsq",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts3",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Ctsm",complete=TRUE)},
\cref{tab:top50_mouse_placenta,fig:cathespin_evolution_tree}).
Finally, \textit{Peg10}, an imprinted gene required for proper
formation of the placenta in mice is also highly expressed
(\Sexpr{gene.expression("mus_musculus","Peg10")}), as previously
reported \cite{Ono.ea2006:DeletionofPeg10imprinted}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Equus Caballus} are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of the annexin complex
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.
